---
title: "Fixed Income Homework 4"
author: "Nitish Ramkumar, Carlos Quicazan, Justin Ge, Yuying Wang"
output: pdf_document
---

```{r echo=FALSE}
library(knitr)
data <- matrix(nrow=5,ncol=6)
data[1,] <- c(0.049,0.058,0.071,0.068,0.075,0.092)
data[2,] <- c(0.049,0.044,0.037,0.056,0.062,0.069)
data[3,] <- c(0.049,0.031,0.026,0.051,0.067,0.048)
data[4,] <- c(0.049,0.05,0.061,0.069,0.069,0.06)
data[5,] <- c(0.049,0.044,0.055,0.061,0.058,0.065)
```

#Question 1  
The duration can be calculated as $e^{-AT}$, where A is average rate for a time horizon T. The duration is as below:    
```{r echo=FALSE}
meanRates <- sapply(2:ncol(data),function(x){
  apply(data[,c(1:x)],1,mean)
})

duration <- sapply(1:ncol(meanRates),function(x){
  exp(-x*meanRates[,x])
})

duration_mean <- apply(duration,2,mean)
kable(t(duration_mean),col.names = c(1:5))
```

#Question 2
Assuming that payoff for the cap at each caplet is the max(interest rate - strike,0) * 100, The mean price of the 5 year interest rate cap is as below:  
```{r echo=FALSE}
strike <- 0.045
payoff <- pmax(data[,-1]-strike,0)*100
pvpayoff <- duration*payoff
price <- apply(pvpayoff,1,sum) 
mean_price <- mean(price)
mean_price
```


#Question 3 
Assuming that payoff for the cap at each floorley is the max(strike - interest rate,0) * 100, The mean price of the 5 year interest rate floor is as below:
```{r echo=FALSE}
strike <- 0.067
payoff <- pmax(strike-data[,-1],0)*100
pvpayoff <- duration*payoff
price <- apply(pvpayoff,1,sum) 
mean_price <- mean(price)
mean_price
```

#Question 4
The price of the 5 year caplet (assuming payoff is on an amount of 100) is as below:
```{r echo=FALSE}
strike <- 0.067
period <- 5
payoff_cap <- pmax(data[,period+1]-strike,0)*100
price_cap <- payoff_cap * duration[,period]
price_cap_mean <- mean(price_cap)
price_cap_mean
```

The price of the 5 year floorlet (assuming payoff is on an amount of 100) is as below
```{r echo=FALSE}
payoff_floor <- pmax(strike - data[,period+1],0)*100
price_floor <- payoff_floor * duration[,period]
price_floor_mean <- mean(price_floor)
price_floor_mean
```
The 5 year Floorlet is more valuable than the 5 year caplet.  
  
#Question 5
The price of the 5 year caplet (assuming payoff is on an amount of 100) is as below:  
```{r echo=FALSE}
strike <- 0.063
period <- 5

payoff_cap <- pmax(data[,period+1]-strike,0)*100
price_cap <- payoff_cap * duration[,period]
price_cap_mean <- mean(price_cap)
price_cap_mean
```

The price of the caplet on 5 year average rate (assuming payoff is on an amount of 100) is as below:  
```{r echo=FALSE}
payoff_avg <- pmax(meanRates[,period]-strike,0)*100
price_avg <- payoff_avg * duration[,period]
price_avg_mean <- mean(price_avg)
price_avg_mean
```

So the 5 year caplet is more valuable than the 5 year caplet on average rates.  

#Question 6
The standard deviation on the 5 year interest rates is as below:
```{r echo=FALSE}
rate5 <- sd(data[,period+1])
rate5
```

The standard deviation on the average interest rates over 5 year period is as below:
```{r echo=FALSE}
avgrate5 <- sd(meanRates[,period])
avgrate5
```
We are more than 99\% confident of the price of the 5 year average rate call option, but we can't be 99\% sure of the price of the 5 year call option    
  
 

