---
title: "Fixed Income Assignment 8"
author: "Nitish Ramkumar, Justin Ge, Carlos Quicazan, Yuying Wang"
output: pdf_document
---

```{r}
library(knitr)
setwd("C:/_UCLA/Quarter3/237F_FixedIncome/Assignment/Assignment8")
corr_mat <- read.csv("corrin.csv",header=FALSE)
corchol <- read.csv("corchol.csv",header=FALSE)
dt_raw <- read.csv("pfilea.csv",header=FALSE)
vol_raw <- read.csv("sigma.csv",header=FALSE)
dt_val <- dt_raw[1:20,]
vol <- vol_raw[1:19,]

set.seed(1234)
noOfSims <- 1000


####Carlos Code
corrin = as.matrix(read.csv(file ="corrin.csv",header=F, sep=","))
corchol = as.matrix(read.csv(file ="corchol.csv",header=F, sep=","))
pfilea = as.matrix(read.csv(file ="pfilea.csv",header=F, sep=","))
sigma = as.matrix(read.csv(file ="sigma.csv",header=F, sep=","))
pfilea = pfilea[1:20]
sigma = sigma[1:20]
sigma_i = sigma[-1]
rho = corrin
dt = 1/2
string_model= function(s,r,D){
  n = length(D)
  D_matrix = cbind(c(D[1],D[1:(n-1)]),matrix(0,ncol = n,nrow = n))
  for (i in 1:(n-1)){
    aux = max(n-i,1)
    sigma_i = sigma_i[1:aux];
    r = r[1:aux,1:aux]
    if (i <= (n-1))
    {
      D_matrix[i,i+1] = 1
      rt = (1/D_matrix[i,i]-1)*2
      Z = rnorm(aux)
      L = t(chol(r))
      dZ = (L %*% Z)*sqrt(dt)
      D_matrix[(i+1):n,i+1] = D_matrix[(i+1):n,i] + D_matrix[(i+1):n,i]*rt*dt+sigma_i*dt*dZ
    }
    if (i == (n-1)) { D_matrix[i+1,i+2] = 1 }
  }
  return(D_matrix)
}
s = sigma[-1]
r = rho[1:19,1:19]
r = rho[-1,-1]
result = array( dim = c(length(pfilea), (length(pfilea)+1), noOfSims))
for (i in 1:noOfSims){
  result[,,i] = string_model(s,r,pfilea)
}

all_sims <- as.list(rep(0,noOfSims))
for(i in c(1:noOfSims)){
  all_sims[[i]] <- result[,,i]
}
####

```

#Question 1

1) Get CMS rates using par rate formula
2) Calculate spot rates at various points for all iterations
3) Get payoff using spot rates and CMS rates (as strike)
4) Discount the payoff back to time 0

```{r}
#use dt provided in excel sheet to calculate cms
cms <- sapply(c(1:length(dt_val)), function(m){2 * 100 * (1 - dt_val[m])/(sum(dt_val[1:m]))})/100

calculateDiscountedPayoffs <- function(all_sims,cms){
  #Seperate All dts for moving one step forward 
  all_dis_fac <- sapply(all_sims,function(x){diag(x)})
  
  #L(t-0.5) is the forward rate at t-0.5, 0.5 steps forward (i.e. until t)
  #a.k.a spot at that moment
  forward_rates <- 2*((1/all_dis_fac)-1)
  
  #Calculate payoff of caplets or swaptions
  payoff <- 0.5*sapply(c(1:length(all_sims)),function(count){pmax(forward_rates[,count]-cms
                                                                  ,0.0)})
  
  #Calculate present value of caplet
  dis_cf <- apply(all_dis_fac,2,cumprod)  
  discount_payoffs <- payoff*dis_cf
  return(discount_payoffs)
}

discounted_payoffs <- calculateDiscountedPayoffs(all_sims,cms)
answer_count <- c(2,3,4,5,7,10)
answers <- sapply(2*answer_count,function(x){mean(apply(discounted_payoffs[2:x,],2,sum))})
kable(cbind(answer_count,answers),col.names = c("cap count","price"))

```

#Question 2

1) For every path, check if at exercise step, rate is greater than strike. 
2) If it is, generate paths from this point (using the string pattern logic, but different initial dt values)  
3) Calculate payoff at each possible step in these newly generated path. This is the fixed leg - floating leg price. Floating leg price = 1. fixed leg is current value of all coupon payments until expiry
4) Discount each payoff using a discount factor
5) Add up cashflows according to what swaption is requested.  


```{r}
calculateSwaptionPrice <- function(all_sims,coupon,exercisestep,expirystep){
  all_dis_fac <- sapply(all_sims,function(x){diag(x)})
  forward_rates <- 2*((1/all_dis_fac)-1)
  
  exercise_val <- forward_rates[exercisestep,]
  itm <- exercise_val > coupon
  payoffs <- rep(0,length(exercise_val))
  for(count in which(itm)){
      dts <-  all_sims[[count]][-c(1:(exercisestep-1)),exercisestep]
      noOfSim2 <- 10
      all_sims2 <- as.list(rep(0,noOfSim2))
      for(incount in c(1:noOfSim2)){
        all_sims2[[incount]] <- string_model(s,r,dts)
      }
      
      #Get present value of fixed rate
      all_dis_fac_2 <- sapply(all_sims2,function(x){diag(x)})
      dis_cf_2 <- apply(all_dis_fac_2,2,cumprod)  
      fixedcf_pv <- coupon * dis_cf_2[-1,]
      total_coupon <- apply(fixedcf_pv[1:expirystep,],2,sum)
      coupon_principal <- total_coupon + dis_cf_2[expirystep,] 
      discount_payoffs <- coupon_principal - 1
      
      payoffs[count] <- mean(discount_payoffs)
  }
  mean(payoffs)
}

n <- 2
expiry <- 2*c(1:4)
forwards_1 <- sapply(expiry, function(m){2 * 100 * 
    (dt_val[n] - dt_val[n+m])/(sum(dt_val[(n+1):(n+m)]))})/100
swaption_price1 <- sapply(c(1:length(forwards_1)),function(x)
                   {calculateSwaptionPrice(all_sims,forwards_1[x],n,expiry[x])})
kable(cbind(c(1:4),swaption_price1),col.names = c("SwpCnt","Price"))

n <- 4
expiry <- 2*c(1:3)
forwards_2 <- sapply(expiry, function(m){2 * 100 * 
    (dt_val[n] - dt_val[n+m])/(sum(dt_val[(n+1):(n+m)]))})/100
swaption_price2 <- sapply(c(1:length(forwards_2)),function(x)
                  {calculateSwaptionPrice(all_sims,forwards_2[x],n,expiry[x])})
kable(cbind(c(1:3),swaption_price2),col.names = c("SwpCnt","Price"))

n <- 10
expiry <- 2*c(1,2,5)
forwards_3 <- sapply(expiry, function(m){2 * 100 * 
    (dt_val[n] - dt_val[n+m])/(sum(dt_val[(n+1):(n+m)]))})/100
swaption_price3 <- sapply(c(1:length(forwards_3)),function(x)
                  {calculateSwaptionPrice(all_sims,forwards_3[x],n,expiry[x])})
kable(cbind(c(1:3),swaption_price3),col.names = c("SwpCnt","Price"))
```


#Question 3

1) For every swaption, for every spot maturity (t = 0.5,1,...20), bump the spot rate by 0.0001 on both sides and use that to find DV01.  
2) Calculation of Dv01 is performed by using the price swaption function from the previous question.  
```{r}

calculateDV01 <- function(pfilea, coupon, exercisestep, expirystep){
  answers <- c()
  for(count in 1:length(pfilea)){
    pfilea_up <- pfilea
    pfilea_down <- pfilea
    rt_up = (1/(pfilea_up[count]^count)-1)*2 + 0.0001
    rt_down = (1/pfilea_down[count]^count-1)*2 - 0.0001
    pfilea_up[count] <- (1 + rt_up/2)^(-count)
    pfilea_down[count] <- (1 + rt_down/2)^(-count)
    
    noOfSims <- 10
    all_sims_up <- as.list(rep(0,noOfSims))
    for (i in 1:noOfSims){
      all_sims_up[[i]] <- string_model(s,r,pfilea_up)
    }
    price_up <- calculateSwaptionPrice(all_sims_up,coupon,exercisestep,expirystep)
    
    all_sims_down <- as.list(rep(0,noOfSims))
    for (i in 1:noOfSims){
      all_sims_down[[i]] <- string_model(s,r,pfilea_down)
    }
    price_down <- calculateSwaptionPrice(all_sims_down,coupon,exercisestep,expirystep)
    
    #answers[count,1] <- count
    answers[count] <- (price_down - price_up)/2
  }
  return(answers)
}

n <- 2
expiry <- 2*c(1:4)
dv01_1 <- sapply(c(1:length(forwards_1)),function(x){calculateDV01(pfilea, forwards_1[x], 
                                                                   n, expiry[x])})

kable(cbind(c(1:20),dv01_1),col.names=c("bump","Swptn1","Swptn2","Swptn3","Swptn4"))

n <- 4
expiry <- 2*c(1:3)
dv01_2 <- sapply(c(1:length(forwards_2)),function(x){calculateDV01(pfilea, forwards_2[x], 
                                                                   n, expiry[x])})

kable(cbind(c(1:20),dv01_2),col.names=c("bump","Swptn5","Swptn6","Swptn7"))

n <- 10
expiry <- 2*c(1,2,5)
dv01_3 <- sapply(c(1:length(forwards_3)),function(x){calculateDV01(pfilea, forwards_3[x], 
                                                                   n, expiry[x])})

kable(cbind(c(1:20),dv01_3),col.names=c("bump","Swptn8","Swptn9","Swptn10"))


matplot(x=c(1:20),y=cbind(dv01_1,dv01_2,dv01_3),type="l",
        col=c("Red","Blue","Green","Black","Yellow","Orange",
              "Brown","Pink","Magenta","Gold"),lty=1,
        main="Plot of DV01 of all the swaptions",xlab="Bump Maturity",ylab="Dv01")
```

#Question 4
1) Use the simulated interest rate paths, to find the payoff, which is current rate on every path - previous rate on that path.  Once we do that, we get cashflow at every step in every path.  
2) Discount these cashflows  
3) Add all the cashflows necessary for the 5 year cap (which is cashflow 2 to 10).

```{r}
calculateResetPayoffs <- function(all_sims){
  #Seperate All dts for moving one step forward 
  all_dis_fac <- sapply(all_sims,function(x){diag(x)})
  
  #L(t-0.5) is the forward rate at t-0.5, 0.5 steps forward (i.e. until t)
  #a.k.a spot at that moment
  forward_rates <- 2*((1/all_dis_fac)-1)
  
  #Calculate payoff of caplets
  payoff <- 0.5*sapply(c(1:length(all_sims)),function(count){
    pmax(forward_rates[2:nrow(forward_rates),count]-
           forward_rates[(1:nrow(forward_rates)-1),count],0.0)})
  
  #Calculate present value of caplet
  dis_cf <- apply(all_dis_fac[2:nrow(all_dis_fac),],2,cumprod)  
  discount_payoffs <- payoff*dis_cf
  return(discount_payoffs)
}

reset_discount_payoffs <- calculateResetPayoffs(all_sims)
#only 1-9, ignoring of first already performed
price <- mean(apply(discounted_payoffs[1:9,],2,sum))
price
```

#Question 5
1) Calculate the cms rate for every maturity for every simulation.  
2) Use this to subtract the strike and to get payoff.
3) Discount the values to t=0 and add up the payoffs

```{r}
calculateCmsPayoffs <- function(all_sims,strike){
  #Seperate All dts for moving one step forward 
  all_dis_fac <- sapply(all_sims,function(x){diag(x)})
  dis_cf <- apply(all_dis_fac,2,cumprod)  
  
  #L(t-0.5) is the forward rate at t-0.5, 0.5 steps forward (i.e. until t)
  #a.k.a spot at that moment
  #forward_rates <- 2*((1/all_dis_fac)-1)
  cms_rates <- t(sapply(c(2:nrow(dis_cf)), function(m){2 * 100 * (1 -dis_cf[m,])/
                                                    (apply(dis_cf[c(1:m),],2,sum))})/100)
  #Calculate payoff of caplets
  payoff <- 0.5*sapply(c(1:length(all_sims)),function(count){
    pmax(cms_rates[,count]-strike,0.0)})
  
  #Calculate present value of caplet
  dis_cf <- apply(all_dis_fac[2:nrow(all_dis_fac),],2,cumprod)  
  discount_payoffs <- payoff*dis_cf
  return(discount_payoffs)
}

cms_payoffs <- calculateCmsPayoffs(all_sims,0.05)
#only 1-9, ignoring of first already performed
price <- mean(apply(cms_payoffs[1:9,],2,sum))
price
```

