---
title: "Computational Homework 6"
author: "Nitish Ramkumar"
output: pdf_document
---

#Question 1

```{r echo=FALSE}
library(knitr)
vol <- seq(0.12,0.48,0.04)
callPrices <- read.csv("Project6/callPrices.csv",header = FALSE)
putPrices <- read.csv("Project6/putPrices.csv",header = FALSE)
prices <- cbind(vol,callPrices,putPrices)
kable(prices,col.names = c("Volatility","Call Prices","Put Prices"),caption="Lookback Prices")
plot(x=prices[,1],y=prices[,2],type="l",main="Lookback prices",ylab="Prices",xlab="Volatility")
lines(x=prices[,1],y=prices[,3],col="red")
legend("bottomright",c("Lookback Call","Lookback Put"),col=c("black","red"),lty=1)
```

#Question 2
##Part a  
  
The default information for the default case are as below: (Default time in years)

```{r echo=FALSE}
allPrices <- read.csv("Project6/AllAnswers.csv",header=FALSE)
allPrices <- allPrices[,c(1:6)]
colnames(allPrices) <- c("lambda1","lambda2","Time","OptionPrice","DefaultTime","Probability")
qa <- unique(allPrices[allPrices$lambda1==0.2 & allPrices$lambda2==0.4 & allPrices$Time==5,])
row.names(qa) <- NULL
kable(qa,col.names=c("Lambda1","Lambda2","Time(years)","Default Option Price","Expected Exercise Time","Probability"))

```

##Part b  
  
As per the question, there will be 6 graphs. 2 each for Default Option Price, Probability and Expected Exercise time.    
Within the 2,  
1) Graph 1 will keep ${\lambda}_1$ constant at 0.2. The graph is a function of time and will have multiple lines for each ${\lambda}_2$.  
2) Graph 2 will keep ${\lambda}_2$ constant at 0.4. The graph is a function of time and will have multiple lines for each ${\lambda}_1$.  
  
  

###Default Option Price
```{r echo=FALSE}
#fix lambda1
fix.lambda1 <- unique(allPrices[allPrices$lambda1 == 0.2,])
fix.lambda2 <- unique(allPrices[allPrices$lambda2 == 0.4,])

colors <- c("black","blue","red","yellow","green","brown","pink","purple","orange")
par(mfrow=c(1,2))
x11(width=2, height=6)
#a - Option Price
plot(y=fix.lambda1[fix.lambda1$lambda2 == 0.0,"OptionPrice"],x=fix.lambda1[fix.lambda1$lambda2 == 0.0,"Time"],main="Option price, lambda1=0.2",xlab="Time",ylab="OptionPrice",type="l",ylim=c(1000,5000))
colCount <- 1
for(val in seq(0.0,0.8,0.1)){
  lines(y=fix.lambda1[fix.lambda1$lambda2 == val,"OptionPrice"],x=fix.lambda1[fix.lambda1$lambda2 == val,"Time"],col=colors[colCount])
  colCount <-  colCount + 1
}
legend("bottomright",as.character(seq(0.0,0.8,0.1)),col=colors[1:colCount-1],lty=1,pt.cex=1,cex=0.5)

plot(y=fix.lambda2[fix.lambda2$lambda1 == 0.05,"OptionPrice"],x=fix.lambda2[fix.lambda2$lambda1 == 0.05,"Time"],main="Option price, lambda2=0.4",xlab="Time",ylab="OptionPrice",type="l",ylim=c(3000,5500))
colCount <- 1
for(val in seq(0.05,0.4,0.05)){
  lines(y=fix.lambda2[fix.lambda2$lambda1 == val,"OptionPrice"],x=fix.lambda2[fix.lambda2$lambda1 == val,"Time"],col=colors[colCount])
  colCount <- colCount + 1
}
legend("bottomright",as.character(seq(0.05,0.4,0.05)),col=colors[1:colCount-1],lty=1,pt.cex=1,cex=0.5)
``` 
  
###Probability
```{r echo=FALSE}
par(mfrow=c(1,2))
x11(width=2, height=4)
plot(y=fix.lambda1[fix.lambda1$lambda2 == 0.0,"Probability"],x=fix.lambda1[fix.lambda1$lambda2 == 0.0,"Time"],main="Probability, lambda1=0.2",xlab="Time",ylab="Probability",type="l",ylim=c(0.2,1))
colCount <- 1
for(val in seq(0.0,0.8,0.1)){
  lines(y=fix.lambda1[fix.lambda1$lambda2 == val,"Probability"],x=fix.lambda1[fix.lambda1$lambda2 == val,"Time"],col=colors[colCount])
  colCount <-  colCount + 1
}
legend("bottomright",as.character(seq(0.0,0.8,0.1)),col=colors[1:colCount-1],lty=1,pt.cex=1,cex=0.4)

plot(y=fix.lambda2[fix.lambda2$lambda1 == 0.05,"Probability"],x=fix.lambda2[fix.lambda2$lambda1 == 0.05,"Time"],main="Probability, lambda2=0.4",xlab="Time",ylab="Probability",type="l",ylim=c(0.73,1))
colCount <- 1
for(val in seq(0.05,0.4,0.05)){
  lines(y=fix.lambda2[fix.lambda2$lambda1 == val,"Probability"],x=fix.lambda2[fix.lambda2$lambda1 == val,"Time"],col=colors[colCount])
  colCount <- colCount + 1
}
legend("bottomright",as.character(seq(0.05,0.4,0.05)),col=colors[1:colCount-1],lty=1,pt.cex=1,cex=0.4)
```
  
###Expected Exercise Time
```{r echo=FALSE}
par(mfrow=c(1,2))
x11(width=2, height=5)
plot(y=fix.lambda1[fix.lambda1$lambda2 == 0.1,"DefaultTime"],x=fix.lambda1[fix.lambda1$lambda2 == 0.1,"Time"],main="Default Time,lambda1=0.2",xlab="Time",ylab="DefaultTime",type="l",ylim=c(0.5,1.7))
colCount <- 1
for(val in seq(0.1,0.8,0.1)){
  lines(y=fix.lambda1[fix.lambda1$lambda2 == val,"DefaultTime"],x=fix.lambda1[fix.lambda1$lambda2 == val,"Time"],col=colors[colCount])
  colCount <-  colCount + 1
}
legend("bottomright",as.character(seq(0.1,0.8,0.1)),col=colors[1:colCount-1],lty=1,pt.cex=1,cex=0.4)

plot(y=fix.lambda2[fix.lambda2$lambda1 == 0.05,"DefaultTime"],x=fix.lambda2[fix.lambda2$lambda1 == 0.05,"Time"],main="Default Time,lambda2=0.4",xlab="Time",ylab="DefaultTime",type="l",ylim=c(0.8,1.5))
colCount <- 1
for(val in seq(0.05,0.4,0.05)){
  lines(y=fix.lambda2[fix.lambda2$lambda1 == val,"DefaultTime"],x=fix.lambda2[fix.lambda2$lambda1 == val,"Time"],col=colors[colCount])
  colCount <- colCount + 1
}
legend("bottomright",as.character(seq(0.05,0.4,0.05)),col=colors[1:colCount-1],lty=1,pt.cex=1,cex=0.4)
